---
title: "Combining Cyclistic Data (July 2024 - June 2025)"
author: "Lyncold Stephensky CHERY"
date: "2025-08-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Install required packages if not already installed
```{r }
packages <- c("tidyverse", "janitor", "skimr")
installed <- packages %in% rownames(installed.packages())
if (any(!installed)) install.packages(packages[!installed])
```

### Load libraries
```{r }
library(tidyverse)
library(janitor)
library(skimr)
```

### Set the working directory to the folder containing raw CSV files
```{r }
setwd("C:/Users/user/Desktop/Cyclistic_Case_Study/data/csv_brut")
```

### List all CSV files in the folder
```{r }
csv_files <- list.files(pattern = "\\.csv$")
```

### Read and combine all CSV files into one dataframe
```{r }
cyclistic_df <- csv_files %>%
  map_df(read_csv)
```

### Display the structure of the combined dataset
```{r }
str(cyclistic_df)
```

### Display a statistical summary of the dataset
```{r }
skim(cyclistic_df)
```

### Save the combined dataset to the cleaned data folder
```{r }
write_csv(cyclistic_df, "Data_Cyclistic_Jul24_to_Jun25.csv")
```

### Delete the raw CSV files if necessary
```{r }
file.remove(csv_files)
```
